### 2011-04-07 Text/paper.tex
Added the introductions. Need to prepare 3D graphs using Sketch3D.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\begin{document}
\begin{abstract}
Classical Bretherton problem describes the propagation of gas fingers through liquid media with
thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
\cite{kuzmin-binary2d}. This
work is the continuation of the previous work. It is focused on the three-dimensional capillaries
simulations with rectangular and square crosssections to
validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
number $0.1\leq Ca \leq 1.0$.  
The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. The results show that the
binary liquid model is able to capture a number of phenomena happening in three-dimensional
capillaries , as the existance of the vortex in front of the bubble and the bubble radii dependency
on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
used to simulate the Bretherton problem with good accuracy.
\end{abstract}

@@

In comparison with the two-dimensional Bretherton flow (circular tubes, flow between parallel
plates) covered in our previous work \cite{kuzmin-binary2d}, there is a vast number of experimental
results available for the three-dimensional case as microchannels with square, triangular and
rectangular crosssections. For instance, \citet{cerro-bubble-train} performed a number of
experiments for a
bubble-train flow in capillaries of
circular and square cross section for horizontal, upward and downward flows. They found a
significant difference in mass transfer between upward and downward flows for the small capillary
number region.

\citet{shikazono-square} obtained
experimental
results for the deposition length dependency on the
capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
shaped mirochannels. They also found experimental correlation for the prediction of bubble radii
based on the capillary number $Ca$ and the Weber number $We=Re Ca$. Thus, there is the effect of
inertia on the bubble formation. It is indicated by \citet{gupta-review} that in comparison with
one phase flows the flows in microchannels depend on the complicated interplay of the gravitational,
interfacial, inertial and viscous forces.
{\color{red} Historically,  it was indicated \cite{bretherton} that inertia effects can be 
neglicted. Later on, the results of \citet{giavedoni-numerical} and
\citet{heil-bretherton} show
{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
range of Reynolds
numbers. 
For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
number from $0$ to $70$ for the flow between plates. Later on, \citet{heil-bretherton} extended
results for the flow between plates upto Reynolds number
$300$. The indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
significantly the pressure distrubution and the flow field near the front bubble tip. Recent
experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
following dependency on the capillary number $Ca$ and the Weber number $We$:
\begin{equation}
\begin{aligned}
&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
&R_{axis}=
\begin{cases}
1, &R_{diag}>1\\
R_{diag}, &R_{diag}\leq 1.
\end{cases}
\end{aligned}
\end{equation}
In the present
simulations the largest Reynolds number we achieve is less than $1$ and the capillary number is of
order of unity. Thus, it allows to neglect 
inertia effects. 
Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
but the
viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
simulation of the square shaped capillaries.

 Moreover, the results based on uniform density, as in the case of the LBM
binary liquid model, are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
gas and liquid, which are of
significantly
different densities and viscosities.
}

The interested reader
is referred to these works for comprehensive experimental correlations.

The experimental works supported by numerical simulations reveal interesting phenomena happening in
three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case,
Fig. \ref{fig:sym:antisym}.
{\color{red} From hereon, we limit ourselves to the case of microchannels with square
crosssections.}
Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. 
{\color{red}
\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
flow in the side of the capillary thus changing the qualitative differencies depending on the
gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
larger fro upflow than for downflow''. 
}
In
this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

\citet{heil-threedim} also indicated another interesting phenomenon. They indicated the transition
capillary number where the streamlines pattern
changes from having a vortex in front of the bubble and not having it for larger capilllary
numbers,
Fig. \ref{fig:streamlines:pattern}.
For the square channel the critical number is $Ca=0.691$.   
%\begin{figure}
%\includegraphics[width=0.5\textwidth]{Figures/ca05_clean.eps}\\
%\includegraphics[width=0.5\textwidth]{Figures/ca10_clean.eps}
%\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
%Streamlines in the moving frame of reference for a tube with $Ca = 0.5$ (top) and $Ca = 2.0$
%(bottom). 
%\label{fig:heil:recirculating}}
%\end{figure}

There are also a number of numerical works for the three-dimensional case. For instance, 
\citet{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and calculated bubble shapes for different
slug and bubble cross sections and menisci appearance.
\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
square and rectangular shaped capillaries. Also, they investigated gravitational effects for
circular shaped capillaries. They also found an empirical correlation which allows to put
different  aspect ratio rectangular channels simulation results on one
curve. \citet{wang-non-circular} performed numerical simulations by the VOF technique for the
non-circular
shaped capillaries as square and triangular shaped capillaries. The VOF technique is the continuous
interface method based on a non-uniform grid. It is of particular interest for us to compare it
with the continuous interface method by the LBM implementation. 
They also performed
numerical simulations for the determination of the critical capillary number $\widehat{Ca}$.

Almost all the simulations indicated above for the two-dimensional and three-dimensional cases
are performed with the help of the finite element method
\cite{ingham-plates,heil-bretherton,heil-threedim}. However, this method is of limited
applicability for
problems
involving complex geometries, free interface motion, or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for this kind of simulations.
However, if the interface is smeared out over several grid nodes, the question of
proper film resolution in comparison with the interface resolution arises.  Therefore, the
\emph{goal} of this work is to reproduce the numerical results and phenomena presented above
with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We
focus on exploring the parameter range of the binary liquid lattice
Boltzmann method for correctly resolving the flow in a range of capillary numbers.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and allows incorporation of
physical phenomena on the microscopic level, as in the case of multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface using continuous interface methods where the interface spans over several grid nodes.
Such a representation
brings issues of the film thickness resolution versus interface
resolution -- the diffuse interface should be dealt with in such a way as to have a
negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
resolution was examined. It was found that the film thickness should be at least as twice as the
interface for the simulation to be accurate.

The binary liquid free-energy LB model due to \citet{swift} we used
simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of
significantly
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects can be 
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
numbers.
Therefore, a major governing factor for microchannel flows is not the density ratio, but the
viscosity ratio. The goal of this work is to do a feasibility study of the LBM
binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
problem. The work results are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton,heil-threedim}.

One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for a relatively large capillary numbers, and
found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}.

The paper is organized as follows.  First, we briefly
explain the binary liquid lattice Boltzmann model. Then, the preliminary results for the
two-dimensional and for the three-dimensional case are presented in the results section. The paper
is
concluded with a summary of the main findings.

\citet{heil-threedim} performed three-dimensional simulations for circular shaped and for the
square shaped capillaries and gravitational effects for circular shaped capillaries. The results for
square capillaries are shown in Fig.\ref{fig:heil:three:dim}. Also, authors performed
numerical simulations for rectangular shaped capillaries. It was found that for different shaped
capillaries there is an empirical correlation allowing to put all the different curves for film
thickness dependency on the capillary number on one curve through introduction of the following
parameter
$s_{\infty}=r_{\infty} \alpha^{-1/2}$. The parameter $r_{\infty}$ is the aximsymmetric bubble radius
infinitely far behind the bubble tip and can be calculated as:
\begin{equation}
r_{\infty}=2\sqrt{(\alpha-Q)/\pi},
\end{equation}
where $Q$ is the net flow rate in the computational domain. {\color{red} What is it? - how to
calculate? Why velocity is involved?}

It is also indicated that for $\alpha\geq\hat{\alpha}=2.04$ the interface can not make up the
axisymmetric case for the range of capillary numbers. The $s_{\infty}$ dependency on $\alpha$ is
shown in Fig.
\ref{fig:heil:three:dim}.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/capillary_width_heil.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/capillary_rectangular_width_heil.eps}\\
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}

\citet{heil-threedim} indicate the interesting occuring thing in square and rectangular channels
as the recirculating regime for flow fields.  It is
known that for certain capillary number the streamlines either go around the bubble or create the
circulating regime in front of the bubble, Fig. \ref{fig:heil:recirculating}.  
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/recirculating.eps}
\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
Streamlines in the moving frame of reference for a rectangular tube $\alpha=1.5$ (a) Ca = 0.2,
(b) Ca = 0.5, (c) Ca = 0.55, (d) Ca = 2.0. Streamlines are shown on the surface of the bubble and
in the plane along the shorter semi-axis.
\label{fig:heil:recirculating}}
\end{figure}
The critical capillary numbers for changing velocity pattern depending on the eccentricity
parameter $\alpha$ are presented in Table \ref{table:recirculation:data}.
\begin{table}
\begin{tabular}{c|c}
$\alpha$& $Ca_{T}$\\
\hline
1& 0.691\\
1.1 & 0.688\\
1.3 & 0.666\\
1.5 & 0.631\\
\end{tabular}
\caption{The results for the recirculation region.\label{table:recirculation:data}}
\end{table}

\citet{shikazono-square} obtained a lot of different results for the deposition length versus the
capillary number variation ethanol/air and water/air experiments. All of them are
presented in Fig. \ref{fig:deposition:han}.
\begin{figure}
%\includegraphics[width=0.47\textwidth]{Figures/han_capillary_fc40_air.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/han_capillary_water_air.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/han_capillary_ethanol_air.eps}\\
\caption{The axial and diagonal radiuses dependency on the capillary number. Courtesy of
\citet{shikazono-square} \label{fig:deposition:han}}
\end{figure}

\citet{cerro-bubble-train} performed a lot of experiments for a buble-train flow in capillaries of
circular and square cross section for horizontal, upward and downward flows. The main results are
presented in Fig. \ref{fig:thulasidas:results}.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/thulasidas_capillary.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/thulasidas_relative_velocity.eps}\\
\caption{\citet{cerro-bubble-train} results for the film thickness
variation and the relative bubble velocity versus the capillary
number.\label{fig:thulasidas:results}}
\end{figure}

\citet{wang-non-circular} performed numerical simulations by VOF technique to for the non-circular
shaped capillaries as square and triangle shaped capillaries. The VOF technique is the continuous
interface method based on the non-uniform grid. It is of particular interest for us to compare it
with the continuous interface method by the LBM implementation. Some interesting results are
obtained for a bubble
shape depending on the capillary number, Fig. \ref{fig:wang:bubble:shape}. They also included
numerical simulations for the determination of the critical capillary number $\hat{Ca}$, Fig.
\ref{fig:wang:critical:capillary}. They also found the relative velocity for the bubble for
different capillary numbers.
\begin{figure}
\includegraphics[width=\textwidth]{Figures/wang_bubble_shape.eps}
\caption{The bubble shape across the length for different capillary numbers.
\label{fig:wang:bubble:shape}}
\end{figure}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/wang_critical_capillary.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/wang_relative_velocity.eps}
\caption{Bubble radiuses variations over the range of capillary numbers (left) and the relative
velocity depending on the capillary number.  \label{fig:wang:critical:capillary}}
\end{figure}

>>>>

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\usepackage{tikz}
\begin{document}
\begin{abstract}
Classical Bretherton problem describes the propagation of gas fingers through liquid media with
thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
\cite{kuzmin-binary2d}. This
work is the continuation of the previous work. It is focused on the three-dimensional capillaries
simulations with rectangular and square crosssections to
validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
number $0.1\leq Ca \leq 1.0$.  
The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. The results show that the
binary liquid model is able to capture a number of phenomena happening in three-dimensional
capillaries, as the existence of the vortex in front of the bubble and the bubble radii dependency
on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
used to simulate the Bretherton problem with good accuracy.
\end{abstract}

@@

The two-dimensional geometry (circular tubes, parallel plates) flows are studied extensively in
experimental works of \citet{aussillous-deposition, cerro-bubble-train} and numerical works of
\citet{giavedoni-numerical,heil-bretherton}. All abovementioned works found that the Bretherton
analysis is valid only for small capillaries numbers $Ca\leq 0.003$ and deviates for larger ones.
That is caused by complex interplay of the gravitational,
interfacial, inertial and viscous forces \cite{gupta-review}. Historically, 
\citet{bretherton} indicated that inertia effects can be 
neglected. \citet{giavedoni-numerical} suggested
that
inertia effects
are negligible for $Ca \leq 0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
number from $0$ to $70$ for two-dimensional geometry flows. Later on, \citet{heil-bretherton}
extended
results for the flow between plates upto Reynolds number
$300$. They indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$), the change in Reynolds number 
influences significantly the pressure distrubution and the flow field near the front bubble tip.
\citet{cerro-bubble-train} show that the mass transfer is strongly affected by flow direction for
small capillary numbers in case of upward and downward flows. Thus, to fully describe the
propagation of the semi-infinite finger in liquid media one needs to take into the account the
viscous, gravitational, surface and inertia forces \cite{gupta-review}.

In comparison with the two-dimensional Bretherton flow  (circular tubes, flow between
parallel
plates), there is a vast number of experimental
results available for the three-dimensional case as microchannels with square, triangular and
rectangular crosssections. For instance, \citet{cerro-bubble-train} performed a number of
experiments for a
bubble-train flow in capillaries of
circular and square cross section for horizontal, upward and downward flows.
\citet{shikazono-square} obtained
experimental
results for the deposition length dependency on the
capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
shaped mirochannels. They also found experimental correlation for the prediction of bubble radii
based on the capillary number $Ca$ and the Weber number $We=Re\,Ca$.

In comparison with two-dimensional geometries, the experimental works
\cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
\cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

There are also a number of numerical works for the three-dimensional case. For instance, 
\citet{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and calculated bubble shapes for different
slug and bubble cross sections and menisci appearance.
\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
square and rectangular shaped capillaries. They indicated the transition
capillary number where the streamlines pattern
changes from having a vortex in front of the bubble and not having it for larger capilllary
numbers, {\color{red} Fig. \ref{fig:streamlines:pattern} }. For the square channel the critical
capillary number is $Ca=0.691$.  Authors also found an empirical correlation which
allows to put
different aspect ratio rectangular channels simulation results on one
curve. It is also indicated that for certain aspect ratio microchannels $\alpha=\frac{a}{b}\geq
2.04$ the interface can not make up the
axisymmetric case for any capillary number. \citet{wang-non-circular} performed numerical
simulations by the VOF technique for the
non-circular
shaped capillaries as square and triangular shaped capillaries. They also investigated the relative
velocity of the bubble for
different capillary numbers.

As it is mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
phenomena are the Volume of Fluid (VOF) method \cite{wang-non-circular}, the level-set method
\cite{fukugata-levelset} and the finite element methods \cite{kreutzer-taylor,heil-threedim}. It is
also indicated \cite{gupta-review}, that the new techniques which are still in the development stage
are the lattice Boltzmann method and phase field methods \cite{anderson-diffuse,gurtin-binary}.
While the finite element methods solve the Bretherton flow as a free-surface problem with a sharp
interface but without underlying details for gas, the lattice Boltzmann method which is a continuous
interface method arises as a promising alternative for simulation of gas finger propagation. The
continuous interface has more flexebility in simulations involving coalescence and/or droplet
breakup.

In our previous work \cite{kuzmin-binary2d} we already covered the Bretherton flow between plates.
It was shown that the free-energy binary liquid model, which is a phase field method, simulates
with reasonable accuracy the Bretherton/Taylor phenomenon. The goal of this work is to examine the
three-dimensional geometry simulations conducted by the free-energy binary liquid lattice Boltzmann
method. We focus on the feasibility study of the numerical method for the microchannels with
square crosssections.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. The method is a particle method which allows to simulate physical
phenomena on the microscopic level. For instance, the introduction of the force or potential on the
microscopic level allows to restore multiphase macroscopic equations \cite{swift,
Shan-chen:extended}.

The binary liquid free-energy LB model due to \citet{swift} we used
simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of
significantly
different densities and viscosities, the parameters were carefully chosen as to neglect the inertia
effects, see Section \ref{sec:numerical:benchmark}. Moreover, the results of
\citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
numbers. Thus, a major governing factor for microchannel flows is not the density ratio, but
the
viscosity ratio. That is why the uniform density binary liquid model is suitable for this kind of
simulations. The goal of this work is to do a feasibility study of the LBM
binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
problem. The work results are in good agreement with other simulations
\cite{heil-threedim,wang-non-circular}.

One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for relatively large capillary numbers, and
found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}. The Shan-Chen model was
used to simulate the Bretherton problem, which is sometimes referred to contain thermodynamically
inconsistent interface \cite{nourgaliev-breakup}.

The paper is organized as follows.  First, we explain the simulation benchmark construction
mainly based on our previous work \cite{kuzmin-binary2d}. Then, the binary liquid lattice
Boltzmann model is outlined. The simulation results for the three-dimensional case are presented in
the results section. The results cover the film thickness dependency on the capillary number, the
relative velocity of the bubble and the vortex profiles. The paper
is concluded with a summary of the main findings.
%The critical capillary numbers for changing velocity pattern depending on the eccentricity
%parameter $\alpha$ are presented in Table \ref{table:recirculation:data}.
%\begin{table}
%\begin{tabular}{c|c}
%$\alpha$& $Ca_{T}$\\
%\hline
%1& 0.691\\
%1.1 & 0.688\\
%1.3 & 0.666\\
%1.5 & 0.631\\
%\end{tabular}
%\caption{The results for the recirculation region.\label{table:recirculation:data}}
%\end{table}

\section{Numerical benchmark approach}
\label{sec:numerical:benchmark}
The main discussion here follows and is based on the two-dimensional benchmark approach
\cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
dimensions. Instead of using

\begin{tikzpicture}
%\draw (-2,-1,0,0)--(0,0,1);
%\draw (0,0,1)--(0,1,1);
\draw (-2,-1,0)--(2,-1,0)--(2,1,0)--(-2,1,0)--cycle;
\draw (-2,-1,0)--(-2,-1,1)--(2,-1,1)--(2,-1,0)--cycle;
\draw (-2,1,0)--(-2,1,1)--(2,1,1)--(2,1,0)--cycle;
\draw (-2,-1,1)--(2,-1,1)--(2,1,1)--(-2,1,1)--cycle;
\draw[rotate=90] (-2,0,0) circle (0.5);
%(Label(XY()*"$x$",align=-3Y));
%//surface s=surface(f,(0,0),(pi,2pi),100,Spline);
%draw("$z$",project(O--X),Arrow);
%draw("$y$",project(O--Y),Arrow);
%draw("$x$",project(O--Z),Arrow);
%//draw(YZ);
%label(project("W",Y,Z,(2,0,-0.3)));
%label(project("L",X,Z,(0,-1,-0.3)));
%label(project("H",Y,Z,(-2,-0.8,0.5)));
%draw(face1);
%draw(face2);
%draw(face3);
%draw(face4);
\end{tikzpicture}
\begin{tikzpicture}[line join=round]
\filldraw[draw=none,fill=white](-2,0)--(2,0)--(2,1)--(-2,1)--cycle;
\draw(2,0)--(2,1)--(-2,1)--(-2,0);
\filldraw[fill=white](-2,-.985)--(2,-.985)--(2,.985)--(-2,.985)--cycle;
\filldraw[fill=white](-2,-.985)--(-2,-1.158)--(2,-1.158)--(2,-.985)--cycle;
\filldraw[draw=none,fill=white](-2,0)--(-2,-1)--(2,-1)--(2,0)--cycle;
\draw(-2,0)--(-2,-1)--(2,-1)--(2,0);
\filldraw[fill=white](-2,1)--(-2,1)--(2,1)--(2,1)--cycle;
\filldraw[draw=none,fill=white](-2,-.087)--(-2,-1.158)--(2,-1.158)--(2,-.087)--cycle;
\draw(-2,-.087)--(-2,-1.158)--(2,-1.158)--(2,-.087);
\filldraw[fill=white](-2,-1)--(-2,-1)--(2,-1)--(2,-1)--cycle;
\filldraw[draw=none,fill=white](-2,.839)--(-2,-1)--(2,-1)--(2,.839)--cycle;
\draw(-2,.839)--(-2,-1)--(2,-1)--(2,.839);
\filldraw[fill=white](-2,.985)--(-2,.811)--(2,.811)--(2,.985)--cycle;
\filldraw[draw=none,fill=white](-2,.839)--(2,.839)--(2,1)--(-2,1)--cycle;
\draw(2,.839)--(2,1)--(-2,1)--(-2,.839);
\filldraw[draw=none,fill=white](-2,-.087)--(2,-.087)--(2,.811)--(-2,.811)--cycle;
\draw(2,-.087)--(2,.811)--(-2,.811)--(-2,-.087);
\end{tikzpicture}% End sketch output

 In order to avoid the mutual

It was indicated that the film thickness should be at least as twice large
as the interface thickness. The interface thickness in the continuous formulation of the
free-energy binary liquid formulation occupies $3-5$ nodes. Therefore, the film should be
resolved at least as $6-10$ nodes.

As it was indicated before \cite{heil-threedim}, the diagonal bubble radius is different from the
axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii coincide for larger values of
$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 a$, where $a$ is the side length of
the square microchannel. Therefore, taking the minimal requirement for the film thickness to be
resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times $ to
simulate

Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. 
{\color{red}
\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
flow in the side of the capillary thus changing the qualitative differencies depending on the
gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
larger fro upflow than for downflow''. 
}

Most multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface using continuous interface methods where the interface spans over several grid nodes.
Such a representation
brings issues of the film thickness resolution versus interface
resolution -- the diffuse interface should be dealt with in such a way as to have a
negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
resolution was examined. It was found that the film thickness should be at least as twice as the
interface for the simulation to be accurate.

Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
but the
viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
simulation of the square shaped capillaries.

 Moreover, the results based on uniform density, as in the case of the LBM
binary liquid model, are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
gas and liquid, which are of
significantly
different densities and viscosities.

Resolving the interface on a fine level for capillary numbers smaller than
$0.003$ is computationally expensive even in the two-dimensional case
(the amount of memory necessary to perform the simulation would be of the order
of tens or hundreds of gigabytes).  This work is therefore based on a
comparison of the results obtained with the known data
validated for the Bretherton bubble flow.  In this paper, we present
techniques for initialization of the simulations and discuss optimal
parameter ranges for the binary liquid LB model, such as the binary liquid parameters, the viscosity
ratio and the grid resolution. The emphasis is to obtain correct flow physics, especially in terms
of film thicknesses.

The VOF technique is the continuous
interface method based on a non-uniform grid. It is of particular interest for us to compare it
with the continuous interface method by the LBM implementation.

Thus, there is the effect of
inertia on the bubble formation.

Recent experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
following dependency on the capillary number $Ca$ and the Weber number $We$:
\begin{equation}
\begin{aligned}
&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
&R_{axis}=
\begin{cases}
1, &R_{diag}>1\\
R_{diag}, &R_{diag}\leq 1.
\end{cases}
\end{aligned}
\end{equation}
In the present
simulations the largest Reynolds number we achieve is less than $1$ and the capillary number is of
order of unity. Thus, it allows to neglect 
inertia effects.
